---
title: "Fig. 2"
author: "Hülsmann & Hartig"
date: "February 15 2018"
output: 
  html_document: 
    fig_height: 7
    keep_md: yes
---

## Load functions and packages

```{r, message = F, warning = F}
library(sads)
library(gnm)
library(MASS)
library(lme4)
library(mgcv)

source("../code/functions_data_simulation.R")
source("../code/functions_analyses.R")
set.seed(123)
```



## Run simulations 

```{r, warning = F}
load(file = "../Data/bciCounts.Rdata")
bciCounts$quadrat = rep(1:5000, times = length(unique(bciCounts$species)))
adult_all = rep(by(bciCounts$A, bciCounts$quadrat, sum), length(unique(bciCounts$species)))
recruit_all = rep(by(bciCounts$B, bciCounts$quadrat, sum), length(unique(bciCounts$species)))
bciCounts$b <- recruit_all - bciCounts$B
bciCounts$a <- adult_all - bciCounts$A

fig2 = runAnalyses(abundanceDist = bci*12, numQuadrats= 5000, dispersal = c(0, 0.4, 0.8, Inf))

fig2[[5]] = runAnalysis(externalData = bciCounts)
names(fig2)[5] = "BCI"

fig2[[6]] = runAnalysis(abundanceDist=bci*12, numQuadrats= 5000, dispersal = 0.1, theta = 1, adultSurvival = "LaManna")
names(fig2)[6] = "LaManna"


```

## Fig. 2

```{r}
gr <- seq(0.8, 0, length.out = 4)
col = c(rgb(gr, gr, gr, 0.25),  rgb(0, 160, 176, alpha = 0.25*255, max = 255), rgb(235, 104, 65, alpha = 0.25*255, max = 255))
col_full = c(rgb(gr, gr, gr, 0.8),  rgb(0, 160, 176, max = 255), rgb(235, 104, 65, max = 255))
pch = c(rep(16,4), 17,15)


oldpar <- par(mfrow = c(2,2), oma = c(3,3,1,1), mar = c(4,4,2,2), las=1, mgp=c(2.5, 0.5, 0), lwd=75/75, cex.axis=3/3, cex.lab=1, font.lab=2, tcl=(-0.2))

for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)[fig2[[j]]$RickerLaManna$CNDD > -10]
  y <- fig2[[j]]$RickerLaManna$CNDD[fig2[[j]]$RickerLaManna$CNDD > -10]
  if (j == 1) {
    plot(x, y, main = "Ricker model", xlim = c(-0.4, 3.4),  ylim = c(-8,2), col = col[1], pch=pch[1], xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k = 3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)

}


for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)
  y <- fig2[[j]]$PowerLaManna$CNDD
  if (j == 1) {
    plot(x, y, main = "Offset-power model", xlim = c(-0.4, 3.4),  ylim = c(-0.5,2.5), col = col[1], pch=pch[1], cex=1, xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
    meanA = seq(min(fig2[[j]]$abundanceA), max(fig2[[j]]$abundanceA), length.out = 20)/10000
    meanB = seq(min(fig2[[j]]$abundanceB), max(fig2[[j]]$abundanceB), length.out = 20)/10000
    estCNDD = log10(meanB + 1)/log10(meanA + 1)
    lines(log10(meanA*50000), estCNDD)
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k=3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
  
}


for(j in 1:6) {
  x <- fig2[[j]]$zerosOnes[fig2[[j]]$RickerLaManna$CNDD > -10]
  y <- fig2[[j]]$RickerLaManna$CNDD[fig2[[j]]$RickerLaManna$CNDD > -10]
  if (j == 1) {
    plot(x, y, main = "",  xlim = c(0,1), ylim = c(-8,2), col = col[1], pch=pch[1], cex=1, xlab = "Proportion recruits without adults", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  if (j != 1) {
    smo <- gam(y ~ s(x, k=3))
    smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
    lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
  }

}


for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)
  y <- fig2[[j]]$PowerRandomR$CNDD
  if (j == 1) {
    plot(x, y, main = "", xlim = c(-0.4, 3.4),  ylim = c(-0.5,2.5), col = col[1], pch=pch[1], cex=1, xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k=3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
}
 
mtext(c("A", "B", "C", "D"), side = rep(3,4), c(-2,-2,-25, -25), at = c(0,0.5, 0, 0.5), outer = T, cex = 1.2, font=2)
```


Fig. 2 Estimated CNDD vs. abundance (Log10 N/ha) per species from the Ricker model (**A**) and the offset-power model (**B**) at the 10x10m scale. Gray circles result from data simulated randomly without CNDD and changing spatial association between adults and recruits, ranging from perfect spatial coupling (lightest gray) to no spatial association (black). Blue triangles depict CNDD estimates for the tropical BCI plot; orange squares the simulation model without CNDD used in the appendix of LaManna *et al*.. For the Ricker model, CNDD bias is highly correlated with the proportion of corrected adult counts (**C**). Fitting a species-specific recruit-to-adult ratio in the offset-power model removes the CNDD-abundance correlation (**D**).



```{r, echo = F}
pdf("Fig2.pdf", width = 8, height = 8)
gr <- seq(0.8, 0, length.out = 4)
col = c(rgb(gr, gr, gr, 0.25),  rgb(0, 160, 176, alpha = 0.25*255, max = 255), rgb(235, 104, 65, alpha = 0.25*255, max = 255))
col_full = c(rgb(gr, gr, gr, 0.8),  rgb(0, 160, 176, max = 255), rgb(235, 104, 65, max = 255))
pch = c(rep(16,4), 17,15)


oldpar <- par(mfrow = c(2,2), oma = c(3,3,1,1), mar = c(4,4,2,2), las=1, mgp=c(2.5, 0.5, 0), lwd=75/75, cex.axis=3/3, cex.lab=1, font.lab=2, tcl=(-0.2))

for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)[fig2[[j]]$RickerLaManna$CNDD > -10]
  y <- fig2[[j]]$RickerLaManna$CNDD[fig2[[j]]$RickerLaManna$CNDD > -10]
  if (j == 1) {
    plot(x, y, main = "Ricker model", xlim = c(-0.4, 3.4),  ylim = c(-8,2), col = col[1], pch=pch[1], xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k = 3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)

}


for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)
  y <- fig2[[j]]$PowerLaManna$CNDD
  if (j == 1) {
    plot(x, y, main = "Offset-power model", xlim = c(-0.4, 3.4),  ylim = c(-0.5,2.5), col = col[1], pch=pch[1], cex=1, xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
    meanA = seq(min(fig2[[j]]$abundanceA), max(fig2[[j]]$abundanceA), length.out = 20)/10000
    meanB = seq(min(fig2[[j]]$abundanceB), max(fig2[[j]]$abundanceB), length.out = 20)/10000
    estCNDD = log10(meanB + 1)/log10(meanA + 1)
    lines(log10(meanA*50000), estCNDD)
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k=3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
  
}


for(j in 1:6) {
  x <- fig2[[j]]$zerosOnes[fig2[[j]]$RickerLaManna$CNDD > -10]
  y <- fig2[[j]]$RickerLaManna$CNDD[fig2[[j]]$RickerLaManna$CNDD > -10]
  if (j == 1) {
    plot(x, y, main = "",  xlim = c(0,1), ylim = c(-8,2), col = col[1], pch=pch[1], cex=1, xlab = "Proportion recruits without adults", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  if (j != 1) {
    smo <- gam(y ~ s(x, k=3))
    smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
    lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
  }

}


for(j in 1:6) {
  x <- log10(fig2[[j]]$abundance)
  y <- fig2[[j]]$PowerRandomR$CNDD
  if (j == 1) {
    plot(x, y, main = "", xlim = c(-0.4, 3.4),  ylim = c(-0.5,2.5), col = col[1], pch=pch[1], cex=1, xlab = "Species abundance log10(N/ha)", ylab = "CNDD")
  } else {
    points(x, y, col = col[j], cex=1, pch=pch[j])    
  }
  smo <- gam(y ~ s(x, k=3))
  smo_fit <- predict(smo, newdata = data.frame(x=sort(unique(x))), type = "response")
  lines(sort(unique(x)), smo_fit, col=col_full[j], lwd=2)
}
 
mtext(c("A", "B", "C", "D"), side = rep(3,4), c(-2,-2,-25, -25), at = c(0,0.5, 0, 0.5), outer = T, cex = 1.2, font=2)
dev.off()
```

