---
title: "Fig. 1"
author: "Hülsmann & Hartig"
date: "February 15 2018"
output: 
  html_document: 
    fig_width: 10
    keep_md: yes
---

## Load functions and packages

```{r, message = F, warning = F}
library(sads)
library(gnm)
library(MASS)
library(lme4)
library(mgcv)

source("../code/functions_data_simulation.R")
source("../code/functions_analyses.R")
set.seed(123)
```


## Run simulations 

```{r, warning = F}
load(file = "../data/bciCounts.Rdata")
bciCounts$quadrat = rep(1:5000, times = length(unique(bciCounts$species)))
adult_all = rep(by(bciCounts$A, bciCounts$quadrat, sum), length(unique(bciCounts$species)))
recruit_all = rep(by(bciCounts$B, bciCounts$quadrat, sum), length(unique(bciCounts$species)))
bciCounts$b <- recruit_all - bciCounts$B
bciCounts$a <- adult_all - bciCounts$A


fig1 = list()

runs = 9

CNDD = seq(0,-2,length.out = runs)
dispersal = seq(0,1,length.out = runs)
adultSurvival = seq(1,0.1,length.out = runs)
suitability = seq(0.1, 1, length.out = runs)
richness = round(10^seq(1,2.6, len = runs))
Aratio = seq(0.05, 0.75, length.out = runs)



vect = list(CNDD, dispersal, adultSurvival, 1-suitability, richness, Aratio)
names = c("CNDD", "Dispersal", "Adult survival", "Habitat specificity", "Species Richness", "Proportion of adults", "BCI")


# CNDD
fig1[[1]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = CNDD, gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd =  rep(0, runs), suitability = rep(1, runs))

# dispersal
fig1[[2]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = dispersal, adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd  = rep(0, runs), suitability = rep(1, runs))

# adult survival
fig1[[3]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = adultSurvival, CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd = rep(0, runs), suitability = rep(1, runs))

# Habitat suitability
fig1[[4]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd = rep(0, runs), suitability = suitability)

# richness
fig1[[5]] = runAnalysesAbundance(which = "ricker", species = richness, dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd  = rep(0, runs), suitability = rep(1, runs))

# Adult ratio
fig1[[6]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = Aratio, ABsd = rep(0, runs), suitability = rep(1, runs))

# true bci
fig1[[length(fig1)+1]] = runAnalysis(which = "ricker", externalData = bciCounts)

```

## Fig. 1

```{r}
par(mfcol = c(2,6), las=1)
op <- par(mfcol = c(2,6), oma = c(2,4,0,0) + 0.1, mar = c(2,0,2,0)+0.3, las=1, mgp=c(1.5, 0.5, 0), lwd=75/75, cex.axis=3/3, cex.lab=2/3, tcl=(-0.2))
col = terrain.colors(runs+1)[1:runs]
for(j in 1:6){
  plot(vect[[j]], fig1[[j]]$summary[,1], xlab = "", ylab = "mean CNDD", ylim = c(-6,0.5), main = names[j], col = NA, bg = col, pch=21, cex=1.6, yaxt="n")
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  plot(NULL, NULL, xlab = "", ylab = "CNDD", xlim = c(-0.4, 3.4), ylim = c(-10,2), yaxt="n")
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  clip(-0.1, 3.1, -100, 100)
  for(i in 1:runs) abline(fig1[[j]]$full[[i]]$lmfit, col = col[i], lwd=1.6)  
  abline(fig1[[length(fig1)]]$lmfit, lwd=1.5)
}
mtext(c("CNDD", "Simulation parameter", "Species abundance log10(N/ha)"), outer = T, side = c(2, 1, 1), las=0, line = c(2, -17.5, -0.2), cex = 0.8, font=2)
mtext(c("A", "B"), side = rep(3,2), c(-4, -32), at = c(-0.04, -0.04), outer = T, cex = 1.2, font=2)

```

Fig. 1 Weighted mean CNDD (**A**) and CNDD-abundance correlations (**B**) estimated by the Ricker model with simulated data, varying ecological parameters (CNDD, dispersal, adult survival, habitat specificity), and community characteristics (species richness, proportion of adults). CNDD is zero except for the CNDD subplots. For details on the simulation settings see (4). Line colors correspond to parameter values in the upper subpanels; results for the Barro Colorado Island (BCI) forest plot are shown in black. Parameters estimation follows LaManna *et al*..