---
title: "Fig. 1"
author: "Lisa Hülsmann & Florian Hartig"
date: "February 15 2018"
output: 
  html_document: 
    fig_width: 10
    keep_md: yes
---


In Fig. 1, we show how various processes (CNDD, dispersal, adult mortality and niche effects) as well as community characteristics (species richness and proportion of adults) affect CNDD estimates of the Ricker model. We varied each parameter and show respective CNDD means and CNDD-abundance correlations. Details on the simulations and parameter settings can be found [here](https://github.com/LisaHuelsmann/CommentTo-LaMannaEtAl-Science/tree/master/code).

*Note that we cannot provide the original dataset in this repository due to the data sharing agreement of the BCI data.* 

It must be requested from the data owners via http://ctfs.si.edu/webatlas/datasets/bci/



## Load packages, functions and BCI data

```{r, message = F, warning = F}
library(sads)
library(gnm)
library(MASS)
library(lme4)
library(mgcv)

source("../code/functions_data_simulation.R")
source("../code/functions_analyses.R")

try(load(file = "../data/bciCounts.Rdata"))
set.seed(123)
```


## Run simulations and analyses

Create settings for processes and community characteristics
```{r, warning = F}
fig1 = list()

runs = 9
CNDD = seq(0,-2,length.out = runs)
dispersal = seq(0,1,length.out = runs)
adultSurvival = seq(1,0.1,length.out = runs)
suitability = seq(0.1, 1, length.out = runs)
richness = round(10^seq(1.1, 2.65, len = runs))
Aratio = seq(0.05, 0.75, length.out = runs)

vect = list(CNDD, dispersal, adultSurvival, 1-suitability, richness, Aratio)
names = c("CNDD", "Dispersal", "Adult survival", "Habitat specificity", "Species Richness", "Proportion of adults", "BCI")
```

Run analyses with simulated data using different CNDD settings
```{r, warning = F}
fig1[[1]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = CNDD, gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd =  rep(0, runs), suitability = rep(1, runs))
```

Run analyses with simulated data using different dispersal settings
```{r, warning = F}
fig1[[2]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = dispersal, adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd  = rep(0, runs), suitability = rep(1, runs))
```

Run analyses with simulated data using different adult survival settings
```{r, warning = F}
fig1[[3]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = adultSurvival, CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd = rep(0, runs), suitability = rep(1, runs))
```

Run analyses with simulated data using different habitat specificity settings
```{r, warning = F}
fig1[[4]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd = rep(0, runs), suitability = suitability)
```

Run analyses with simulated data using different species richness settings
```{r, warning = F}
fig1[[5]] = runAnalysesAbundance(which = "ricker", species = richness, dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = rep(0.4, runs), ABsd  = rep(0, runs), suitability = rep(1, runs))
```

Run analyses with simulated data using different adult ratios
```{r, warning = F}
fig1[[6]] = runAnalysesAbundance(which = "ricker", species = rep(200,runs), dispersal = rep(0.5, runs), adultSurvival = rep(1, runs), CNDD = rep(0, runs), gaps = rep(1, runs), Aratio = Aratio, ABsd = rep(0, runs), suitability = rep(1, runs))
```

Run analyses with real counts from the BCI data
```{r, warning = F}
if(exists("bciCounts")) fig1[[length(fig1)+1]] = runAnalysis(which = "ricker", externalData = bciCounts)
```


## Fig. 1

```{r}
par(mfcol = c(2,6), las=1)
op <- par(mfcol = c(2,6), oma = c(2,4,0,0) + 0.1, mar = c(2,0,2,0)+0.3, las=1, mgp=c(1.5, 0.5, 0), lwd=75/75, cex.axis=3/3, cex.lab=2/3, tcl=(-0.2))
col = terrain.colors(runs+1)[1:runs]
for(j in 1:6){
  plot(vect[[j]], fig1[[j]]$summary[,1], xlab = "", ylab = "mean CNDD", ylim = c(-6,0.5), main = names[j], col = NA, bg = col, pch=21, cex=1.6, yaxt="n", xaxt = "n", log = ifelse(j==5, "x", ""))
  if(j!=5) axis(1)
  if(j==5) axis(1, at = rep(1:2, each=2)*10^(1:2))
  if(j==5) axis(1, at = c(10*1:9, 100*1:9), labels = F)
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  plot(NULL, NULL, xlab = "", ylab = "CNDD", xlim = c(-0.4, 3.4), ylim = c(-10,2), yaxt="n")
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  clip(-0.1, 3.1, -100, 100)
  for(i in 1:runs) abline(fig1[[j]]$full[[i]]$lmfit, col = col[i], lwd=1.6)  
  if(exists("bciCounts")) abline(fig1[[length(fig1)]]$lmfit, lwd=1.5)
}
mtext(c("CNDD", "Simulation parameter", "Species abundance log10(N/ha)"), outer = T, side = c(2, 1, 1), las=0, line = c(2, -17.5, -0.2), cex = 0.8, font=2)
mtext(c("A", "B"), side = rep(3,2), c(-4, -32), at = c(-0.04, -0.04), outer = T, cex = 1.2, font=2)

```

**Fig. 1** Weighted mean CNDD (**A**) and CNDD-abundance correlations (**B**) estimated by the Ricker model with simulated data, varying ecological parameters (CNDD, dispersal, adult survival, habitat specificity), and community characteristics (species richness, proportion of adults) at the 10x10m scale. CNDD is zero except for the CNDD subplots. For details on the simulation settings see [here](https://github.com/LisaHuelsmann/CommentTo-LaMannaEtAl-Science/tree/master/code). Line colors correspond to parameter values in the upper subpanels; black lines depict CNDD-abundance correlations for the tropical Barro Colorado Island (BCI) forest plot. Parameters estimation follows LaManna *et al*..



```{r, echo = F, results=FALSE}
pdf("Fig1.pdf", width = 10, height = 5)
par(mfcol = c(2,6), las=1)
op <- par(mfcol = c(2,6), oma = c(2,4,0,0) + 0.1, mar = c(2,0,2,0)+0.3, las=1, mgp=c(1.5, 0.5, 0), lwd=75/75, cex.axis=3/3, cex.lab=2/3, tcl=(-0.2))
col = terrain.colors(runs+1)[1:runs]
for(j in 1:6){
  plot(vect[[j]], fig1[[j]]$summary[,1], xlab = "", ylab = "mean CNDD", ylim = c(-6,0.5), main = names[j], col = NA, bg = col, pch=21, cex=1.6, yaxt="n", xaxt = "n", log = ifelse(j==5, "x", ""))
  if(j!=5) axis(1)
  if(j==5) axis(1, at = rep(1:2, each=2)*10^(1:2))
  if(j==5) axis(1, at = c(10*1:9, 100*1:9), labels = F)
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  plot(NULL, NULL, xlab = "", ylab = "CNDD", xlim = c(-0.4, 3.4), ylim = c(-10,2), yaxt="n")
  if(j==1) axis(2)
  abline(h = 0, col="grey", lty=2)
  clip(-0.1, 3.1, -100, 100)
  for(i in 1:runs) abline(fig1[[j]]$full[[i]]$lmfit, col = col[i], lwd=1.6)  
  if(exists("bciCounts")) abline(fig1[[length(fig1)]]$lmfit, lwd=1.5)
}
mtext(c("CNDD", "Simulation parameter", "Species abundance log10(N/ha)"), outer = T, side = c(2, 1, 1), las=0, line = c(2, -17.5, -0.2), cex = 0.8, font=2)
mtext(c("A", "B"), side = rep(3,2), c(-4, -32), at = c(-0.04, -0.04), outer = T, cex = 1.2, font=2)
dev.off()
```

